Learning Actionable Analytics (with applications for reducing defects and reducing runtimes)

### 2015-05-03 how.tex
Updates from ShareLaTeX

In summary, this paper makes five contributions:
\be
\item A general framework for augmenting a prediction system such that
it becomes a planning system;
\item A  test that recognizes  when we should {\em not}  plan: in summary, our kind of planning is
not recommended in domains that  generate poorly performing predictors;
\item A general verification rig for assessing the value of plans generated within this rig.
\item The HOW  instance-based planner.
\item A proof that HOW runs in near-linear time.
\item An experimental validation that the HOW planner is useful for reducing the
expected values of software defects and software runtimes (at least, for the data sets studied here).
\ee
The rest of this paepr is structued as follows.

@@

\subsection{Testing for Grainer Causality}
With one small variation,  the above framework can be used to check if the learned
plans can be said to {\em cause} a change in the domain being explored.

In the literature, there are two main definitions of causality: Hill's H-causality
and Grainer's G-casualty. 
Of the two, Hill's definition is the most strict. An effect
is H-causal if it satisfies nine criteria including:
\be
\item
Plausibility;
\item
Strength of the effect; 
\item
Specificity: locality of the effect
to specialized groups;
\item
Support by experimental evidence;
\item
Coherence between cause-effects noted
in the field and in laboratory samples;
\item
Temporally: the effect has to follow the cause;
\item
Analogy: similar causes have similar effects;
\item
Biological gradient: more exposure to the cause leads to more of an
effect)
\item
Consistency: the same cause-effect pairings have been seen by different persons
in different places.
\ee
It is hard to demonstrate H-causality in software. For example, Hill's criteria emphasises repeatability and laboratory studies-- both of which are
hard to perform in software engineering (e.g. see the continued debate about
the value of using undergraduate students to surrogates for industrial
practitioners in software engineering research~\cite{Carver2003}). Also, Hill's requirement
for consistency is easier to achieve in studies of small closed systems (e.g. bacteria in a petri dish)
than in large socio-technical systems (e.g. open source projects). As Lofti Zadeh
commented: ``as complexity rises, precise statements lose meaning and meaningful statements lose precision''~\cite{zadeh.

Granger causality is a statistical concept of causality that is based on prediction. According to Granger causality, if a signal X1 "Granger-causes" (or "G-causes") a signal X2, then past values of X1 should contain information that helps predict X2 above and beyond the information contained in past values of X2 alone. Its mathematical formulation is based on linear regression modeling of stochastic processes (Granger 1969).

Hill1965

\cite{Seth2007,granger80}

@@

% \section{Discussion}
% \section{Threats to validity}
% \section{Conclusion}
% \section*{Acknowledgements}
\newpage
\bibliography{References}{}
\bibliographystyle{IEEEtran}
\end{document}

>>>>

In summary, this paper makes five contributions:
\be
\item A general framework for augmenting a prediction system such that
it becomes a planning system;
\item A  test that recognizes  when we should {\em not}  plan: in summary, our kind of planning is
not recommended in domains that  generate poorly performing predictors;
\item A general verification rig for assessing the value of plans generated within this rig.
\item The HOW  instance-based planner.
\item A proof that HOW runs in near-linear time.
\item An experimental validation that the HOW planner is useful for reducing the
expected values of software defects and software runtimes (at least, for the data sets studied here).
\ee
The rest of this paper is structured as follows.

@@

\subsection{Are These Plans ``Causal''?}
With one small variation,  the above framework can be used to check if the learned
plans can be said to {\em cause} a change in the domain being explored.

In the literature, there are two main definitions of causality: Hill's H-causality~\cite{Hill1965}
and Granger's G-casualty~\cite{granger80}. 
Of the two, Hill's definition is the most strict. An effect
is H-causal if it satisfies nine criteria including:
\be
\item
Plausibility;
\item
Strength of the effect; 
\item
Specificity: locality of the effect
to specialized groups;
\item
Support by experimental evidence;
\item
Coherence between cause-effects noted
in the field and in laboratory samples;
\item
Temporally: the effect has to follow the cause;
\item
Analogy: similar causes have similar effects;
\item
Biological gradient: more exposure to the cause leads to more of an
effect)
\item
Consistency: the same cause-effect pairings have been seen by different persons
in different places.
\ee
It is hard to demonstrate H-causality in software engineering.
For example, Hill's criteria emphasises repeatability and laboratory studies-- both of which are
hard to perform in software engineering (e.g. see the continued debate about
the value of using undergraduate students to surrogates for industrial
practitioners in software engineering research~\cite{Carver2003}). Also, Hill's requirement
for consistency is far easier to achieve in studies of small closed systems (e.g. bacteria in a petri dish)
than in large socio-technical systems (e.g. open source projects).

The practical difficulties associated with Hill's criteria lead Granger~\cite{granger80}
to propose another criteria\footnote{Granger  notes that
his proposal was based on earlier work by Norbert Weiner~\cite{Seth2007}.}.
G-causality is based on building predictions based on {\em past} data,
then showing that this works for {\em future} data~\cite{granger80}.

Note that the standard 10-way cross-validation
procedure  in data mining does {\em not} demonstrate G-causality.  In a N-way cross-val experiments, the data is divided into $N$ bins.  
$N-1$ bins are used for learning and the results are tested
on the  remaining bin.
Cross-validation is then repeated $N$ times, each time using a different bin for the test set.
This approach can not be used to demonstrate causality since, if the initial data is collected
over (say) $N$ months, then it is possible that the test set contains observations collected
{\em before}   items seen in the training set.

Jiang et al.~\cite{me11f} and Lumpe et al.~\cite{me11f} propose a modification to cross-validation which lets them
demonstate G-causal defect effects in softwrare engineering.
They divide project data  into $N$ bins, each of with contain data collected at similar times.
They then sort those bins (on the collection date) and use bins $i,j$ to predict for $k$ where 
$i<j<k$.

We adapt the work of Jiang,Lumpe et al.  as follows. In  \fig{work}, we demand that the {\em test}
data is not collected before the {\em train} data.  With that change,
we can reasonably claim that the plan's generating by HOW are G-causal.

@@

% \section{Discussion}
% \section{Threats to validity}
% \section{Conclusion}
% \section*{Acknowledgements}
\newpage
\bibliography{how}{}
\bibliographystyle{IEEEtran}
\end{document}
